---
output:
  html_document: default
  pdf_document: default
---
Data Analysis with R - Wine Quality White by Michelle_Beckner
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
wqw <- read.csv("wineQualityWhites.csv")
```

The dataset I choose to examine is white wine quality. White wine samples from 
variants of the Portuguese "Vinho Verde" wine were evaluated by wine experts. 
The dataset consists of 11 inputs and one output. The inputs are the following:

   1 - fixed acidity
   2 - volatile acidity 
   3 - citric acid 
   4 - residual sugar
   5 - chlorides 
   6 - free sulfur dioxide 
   7 - total sulfur dioxide 
   8 - density
   9 - pH
   10 - sulfates 
   11 - alcohol 
   
There was 1 output. The output was quality as measured with a score from 1 t0 10 
by the wine experts.

  P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
  Modeling wine preferences by data mining from physicochemical properties.
  In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

  Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016
                [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/
                winequality09.pdf
                [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib


Univariate Plots Section   

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
class(wqw)
```
The data is organized into a data frame. 
                
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots2}
dim(wqw)
```
The data frame has 4898 rows and 13 columns.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots3}
str(wqw)
```

The above chart summarizes what I already found out with the data is in a data 
frame, has 4898 rows and 13 columns. I also see the column names and that the 
columns are made up of integers and numeric (decimals). 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots4}
head(wqw)
```

The first row consists of column names. Each row represents the information for 
one sample of white wine.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots5}

library(gridExtra)
library(ggplot2)

grid.arrange(qplot(x=residual.sugar, data = wqw, xlim =c(0, 21),
      ylab = 'Number of Samples',
      xlab = 'Residual Sugar'),
qplot(x=fixed.acidity, data = wqw, xlim =c(4, 11),
      ylab = 'Number of Samples',
      xlab = 'Fixed Acidity'),
qplot(x=volatile.acidity, data = wqw,
      ylab = 'Number of Samples',
      xlab = 'Volatile Acidity'),
qplot(x=citric.acid, data = wqw, xlim =c(0, 0.95),
      ylab = 'Number of Samples',
      xlab = 'Citric Acid'),
qplot(x=free.sulfur.dioxide, data = wqw, xlim =c(0, 110),
      ylab = 'Number of Samples',
      xlab = 'Free Sulfur Dioxide'),
qplot(x=total.sulfur.dioxide, data = wqw, xlim =c(0, 275),
      ylab = 'Number of Samples',
      xlab = 'Total Sulfur Dioxide'),
qplot(x=density, data = wqw, xlim =c(0.985, 1.002),
      ylab = 'Number of Samples',
      xlab = 'Density'),
qplot(x=pH, data = wqw,
      ylab = 'Number of Samples',
      xlab = 'pH'),
qplot(x=sulphates, data = wqw,
      ylab = 'Number of Samples',
      xlab = 'Sulphates'),
qplot(x=alcohol, data = wqw,
      ylab = 'Number of Samples',
      xlab = 'Alcohol'),
qplot(x=quality, data = wqw,
      ylab = 'Number of Samples',
      xlab = 'Quality'))
```


The above charts show the how many samples use how much of each variable. The 
quality variable is the one that most resembles a standard distribution with 
most of the samples receiving the middle ratings. The residual sugar graph I 
found interesting with it being highly left skewed followed by a long tail. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots6}
qplot(x=residual.sugar, data = wqw, xlim =c(0, 25),
      ylab = 'Number of Samples',
      xlab = 'Residual Sugar')+
    facet_wrap(~quality)
```

I was curious of the residual sugar was differently skewed between the different 
quality ratings. The above charts show the residual sugar broken out by the 
different quality ratings. For the most part, it appears that for the middle 
ratings, it keeps the trend of being left skewed. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots7}
qplot(x=alcohol, data = wqw, xlim =c(7, 14),
      ylab = 'Number of Samples',
      xlab = 'Alcohol')+
    facet_wrap(~quality)
```

The above charts show the alcohol content broken out by the different quality
ratings. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots8}
ggplot(data = wqw, aes(x = quality))+
    geom_histogram(binwidth = 1)+
    scale_x_continuous(breaks = seq(1, 10, 1))+
    theme_set(theme_minimal(12))+
    labs(x = "Quality Rating",  y = "Number of Samples")
```

The wine quality was rated from 1 - 10 with the rating of 6 being given the 
most. There was no sample that received a 1 or a 10 rating.

```{r echo=FALSE, Univariate_Plots9}
table(wqw$quality)
```

The above table shows how many samples were rated in each quality category.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots10}
qplot(x=alcohol, y=density, data = wqw, geom = 'boxplot')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots11}
qplot(x=alcohol, y=residual.sugar, data = wqw, geom = 'boxplot')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots12}
qplot(x=residual.sugar, y=density, data = wqw, geom = 'boxplot')
```

The above 3 box charts just shows some more relationships between the 3
variables that I have been looking at: alcohol, density and residual sugar.

# Bivariate Plots Section

I want to see if there is correlation between alcohol content and density.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots1}
qplot(x = alcohol, y = density, data = wqw)
```

From the above chart, it shows that the lower the alcohol content is the lower 
the higher the density.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots2}
qplot(x = quality, y = density, data = wqw)
```

This chart appears to show the the higher quality scores have lower density. 
Although I do not think this is the best chart type to reflect that. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots3}
cor.test(wqw$alcohol, wqw$density, method = "pearson")
```

The pearson score of 0.78 shows there is a significant correlation between
alcohol and density. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots4}
ggplot(aes(x = alcohol, y = quality), data = wqw) +
    geom_point(alpha = 0.05,
             position = position_jitter(h=0),
             colour = "green") +
    geom_line(stat = "summary", fun.y = mean)
```

The above graph shows quality compared to alcohol content. The mean is also
represented. It appears that the higher the alcohol content the higher rating 
the wine sample tended to receive. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots5}
cor.test(wqw$alcohol, wqw$quality, method = "pearson")
```

The pearson score of 0.43 shows there is a correlation between
alcohol and quality. However, it's not a real strong one. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots6}
ggplot(aes(x = residual.sugar, y = quality), data = wqw) +
  xlim(0, 20) +
  geom_point(alpha = 0.05,
             position = position_jitter(h=0),
             colour = "green") +
    geom_line(stat = "summary", fun.y = mean)
```

The above chart shows a quality ratings with the residual sugar content of the 
samples. It appears that there is a small correlation between the amount of 
residual sugar and the quality ratings of the samples. I will do a pearson test
to see if this is correct.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots7}
cor.test(wqw$residual.sugar, wqw$quality, method = "pearson")
```

With a -0.98 pearson correlation, it validates our throughts that there is not
really a correlation between residual sugar and quality ratings.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(data = wqw,
      aes(y = density, x = alcohol,
           color = quality)) +
      ylim(0.985, 1.002) +
    geom_point(alpha = 0.9, size = 3) 
```

The above chart shows the density and alcohol comparison by adding the quality
rating as well. The brighter the color, the higher the quality rating. This
shows that the higher alcohol content not only has the lower density but 
higher rating. This shows what we have been seeing that the higher the 
alcohol content, the lower the density and the higher the quality rating.

# Final Plots and Summary

### Plot One

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(aes(x = alcohol, y = density), data = wqw) +
    ylim(0.985, 1.002) +
    geom_point(alpha = 0.05,
             position = position_jitter(h=0),
             colour = "green") +
    geom_line(stat = "summary", fun.y = mean) +
    xlab("Alcohol") +
    ylab("Density") +
    ggtitle("Relationship Between the Density and Alcohol Variables")
```

### Description One

The above chart shows the density variable compared to the alcohol variable 
that was explored above. The density mean was added for extra clarification.
The chart further shows that the higher the density the lower the alcohol
content. 


### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
ggplot(aes(x = alcohol, y = quality), data = wqw) +
    geom_point(alpha = 0.05,
             position = position_jitter(h=0),
             colour = "purple") +
    geom_line(stat = "summary", fun.y = mean) +
    xlab("Alcohol") +
    ylab("Quality") +
    ggtitle("Relationship Between the Quality Rating and Alcohol Variable")
```

The above graph shows quality compared to alcohol content. The mean is also
represented. It appears that the higher the alcohol content the higher rating 
the wine sample tended to receive. 

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(data = wqw,
      aes(y = density, x = alcohol,
           color = quality)) +
      ylim(0.985, 1.002) +
    geom_point(alpha = 0.9, size = 3) +
    xlab("Alcohol") +
    ylab("Density") +
    ggtitle("Relationship Between the Quality Rating and the Alcohol and Density 
            Variables")
```

The above brings all three factors together - alcohol content, density and 
quality rating. The brighter the color, the higher the quality rating. This
shows that the higher alcohol content not only has the lower density but 
higher rating. This shows what we have been seeing that the higher the 
alcohol content, the lower the density and the higher the quality rating.
------

# Reflection

This was an interesting look at wine and the different variables that go into
different wines. It wasn't a surprise that the higher the alcohol content
corresponded with the higher quality ratings. What did surprise me was that 
the residual sugar had no real impact on the quality ratings. I also was
not expecting the density to play a factor. However, these charts show that the
higher rated wines had lower density.

#Resources

N/A. 
